---
title: "Homework 8: Fake data"
author: "AWhitten"
date: "5/25/2021"
output: html_document
---
# Data parameters
I chose to calculate my parameters (n, mean, and sd) from an existing one year dataset using bluegill lengths from the three pools that make up the upper Illinois River.  
```{r}
# Library -------------------------------
library(dplyr)
library(ggplot2)
# DataSet -------------------------------
my_data <- read.table(file="C:/Users/Kris/Desktop/Bio381Scripting/blg.csv", header=TRUE,sep=",",comment.char="#") 
str(my_data)

# Extract Data Needed -------------------------------
starvedR<-my_data %>% dplyr:: filter(reach == "I05") 
head(starvedR)
sr_mean <- mean(starvedR$length)
sr_sd <- sd(starvedR$length)
sr_cout <- sum(starvedR$count)

Mars<-my_data %>% dplyr:: filter(reach == "I04") 
head(Mars)
ma_mean <- mean(Mars$length)
ma_sd <- sd(Mars$length)
ma_cout <- sum(Mars$count)

DresdenI<-my_data %>% dplyr:: filter(reach == "I03") 
head(DresdenI)
di_mean <- mean(DresdenI$length)
di_sd <- sd(DresdenI$length)
di_cout <- sum(DresdenI$count)
```

# Fake Data
Using identified parameters create a fake dataset using a random normal distribution looking at the differences in bluegill lengths in the upper three pools of the upper Illinois River. 
```{r}
# Creating Random data-------------------------------
# set for each pool of the upper illinois river 
star<- rnorm(n = 100, mean =sr_mean, sd = sr_sd)
mars <- rnorm(n=100, mean = ma_mean, sd = ma_sd)
dres <- rnorm(n= 100, mean = di_mean, sd = di_sd)
# organize dataframe-------------------------------
upperR <- data.frame(ID=1:300,c(star,mars,dres), Treatment = c(rep("starved",100), rep("marseilles", 100), rep("dresden", 100)))
colnames(upperR) [2] <- ("length")
str(upperR)

# Stats- ANOVA -------------------------------
stats <- aov(length~Treatment, data = upperR)
stat_sum <- summary(stats)
print(stat_sum)

# Results boxplot -------------------------------
ano_plot <-ggplot(stats) + aes(x=Treatment, y=length) + geom_boxplot()
print(ano_plot)
```

# Run Again
Look at differences when run another set of fake data. 
```{r}
########################
# Run Again -Creating Random data-------------------------------
# set for each pool of the upper illinois river 
star<- rnorm(n = 100, mean =sr_mean, sd = sr_sd)
mars <- rnorm(n=100, mean = ma_mean, sd = ma_sd)
dres <- rnorm(n= 100, mean = di_mean, sd = di_sd)
# organize dataframe-------------------------------
upperR <- data.frame(ID=1:300,c(star,mars,dres), Treatment = c(rep("starved",100), rep("marseilles", 100), rep("dresden", 100)))
colnames(upperR) [2] <- ("length")
str(upperR)

# Stats- ANOVA -------------------------------
stats <- aov(length~Treatment, data = upperR)
stat_sum <- summary(stats)
print(stat_sum)

# Results boxplot -------------------------------
ano_plot <-ggplot(stats) + aes(x=Treatment, y=length) + geom_boxplot()
print(ano_plot)
```

# Minimum Sample Size
Manipulate the data to determine what the minimum sample size is to still generate a significant result.
```{r}
########################
# Minimum Sample size -------------------------------
# how small can the sample size be to get a significant difference between treatments?
# Smallest sample size with sig difference was 20/treatment
# Run Again -Creating Random data-------------------------------
# set for each pool of the upper illinois river 
star<- rnorm(n = 20, mean =sr_mean, sd = sr_sd)
mars <- rnorm(n=20, mean = ma_mean, sd = ma_sd)
dres <- rnorm(n= 20, mean = di_mean, sd = di_sd)
# organize dataframe-------------------------------
upperR <- data.frame(ID=1:60,c(star,mars,dres), Treatment = c(rep("starved",20), rep("marseilles", 20), rep("dresden", 20)))
colnames(upperR) [2] <- ("length")
str(upperR)

# Stats- ANOVA -------------------------------
stats <- aov(length~Treatment, data = upperR)
stat_sum <- summary(stats)
print(stat_sum)

# Results boxplot -------------------------------
ano_plot <-ggplot(stats) + aes(x=Treatment, y=length) + geom_boxplot()
print(ano_plot)
```

# Effect Size
Adjust the means to look at effect size
```{r}
########################
# Adjust means- Effect Size -------------------------------
# mean for Dresden has to be at least 91 for the ANOVA to keep its significance. Starved Rock and Marseilles are not different at all.

# Run Again -Creating Random data-------------------------------
# set for each pool of the upper illinois river 
star<- rnorm(n = 100, mean =79, sd = sr_sd) # original 79
mars <- rnorm(n=100, mean = 76, sd = ma_sd) # original 76
dres <- rnorm(n= 100, mean = 91, sd = di_sd) # original 116
# organize dataframe-------------------------------
upperR <- data.frame(ID=1:300,c(star,mars,dres), Treatment = c(rep("starved",100), rep("marseilles", 100), rep("dresden", 100)))
colnames(upperR) [2] <- ("length")
str(upperR)

# Stats- ANOVA -------------------------------
stats <- aov(length~Treatment, data = upperR)
stat_sum <- summary(stats)
print(stat_sum)

# Results boxplot -------------------------------
ano_plot <-ggplot(stats) + aes(x=Treatment, y=length) + geom_boxplot()
print(ano_plot)
```