---
title: "Homework 7"
author: "AWhitten"
date: "4/20/2021"
output: html_document
---
# Simulating and Fitting data Distributions


### Open Libraries
```{r}
library(ggplot2) # for graphics
library(MASS) # for maximum likelihood estimation
```
### Read in data vector
#### To illustrate, generate some fake data here:
````{r}
# quick and dirty, a truncated normal distribution to work on the solution set
# z <- rnorm(n=3000,mean=0.2)
# z <- data.frame(1:3000,z)
# names(z) <- list("ID","myVar")
# z <- z[z$myVar>0,]
# str(z)
# summary(z$myVar)
````
### Read in my sample data
```{r}
# Using 2017 CCF length data for example
z <- read.table("MyDataFile.csv",header=TRUE,sep=",", stringsAsFactors=FALSE) # two variables "ID" and "myvar"
str(z)
summary(z)

```
#### Fit different statistical distributions to a vector of observations and estimate the maximum likelihood parameters for each distribution

### Plot histogram of data
#### Plot a histogram of the data with modified code from lecture. Switch from qplot to ggplot for more graphic options and rescale y axis for the histogram from counts to denisty so the area under the histogram equals 1.0.
````{r}
p1 <- ggplot(data=z, aes(x=myVar, y=..density..)) +
  geom_histogram(color="grey60",fill="cornsilk",size=0.2) 
print(p1)
````


### Add Emperical Density Curve
#### Modify to add in kernel density plot of data. This is an empirical curve that is fitted. It doesn't assume a particular probability distrbution but smooths out histogram shape.
````{r}
p1 <-  p1 +  geom_density(linetype="dotted",size=0.75)
print(p1)
````

### Get maximum likelihood paramters for normal distribution
#### Fit normal distribution to data and grab max likelihood estimators of the two parameters for normal (mean and variance)
````{r}
normPars <- fitdistr(z$myVar,"normal")
print(normPars)
str(normPars)
normPars$estimate["mean"] # note structure of getting a named attribute
````

### Plot normal probability density
#### Call the dnorm function inside ggplot’s stat_function to generate the probability density for the normal distribution. Read about stat_function in the help system to see how you can use this to add a smooth function to any ggplot. Note that we first get the maximum likelihood parameters for a normal distribution fitted to thse data by calling fitdistr. Then we pass those parameters (meanML and sdML to stat_function:
````{r}
meanML <- normPars$estimate["mean"]
sdML <- normPars$estimate["sd"]

xval <- seq(0,max(z$myVar),len=length(z$myVar))

 stat <- stat_function(aes(x = xval, y = ..y..), fun = dnorm, colour="red", n = length(z$myVar), args = list(mean = meanML, sd = sdML))
 p1 + stat
````
#### Notice for fake data that the best-fitting normal distribution (red curve) for these data actually has a biased mean. That is because the data set has no negative values, so the normal distribution (which is symmetric) is not working well.


### Plot Exponential Probability Density
#### Use the same template and add in the curve for the exponential
````{r}
expoPars <- fitdistr(z$myVar,"exponential")
rateML <- expoPars$estimate["rate"]

stat2 <- stat_function(aes(x = xval, y = ..y..), fun = dexp, colour="blue", n = length(z$myVar), args = list(rate=rateML))
 p1 + stat + stat2
````

### Plot Uniform Probability Density
##### For the uniform, we don’t need to use fitdistr because the maximum likelihood estimators of the two parameters are just the minimum and the maximum of the data:
````{r}
stat3 <- stat_function(aes(x = xval, y = ..y..), fun = dunif, colour="darkgreen", n = length(z$myVar), args = list(min=min(z$myVar), max=max(z$myVar)))
 p1 + stat + stat2 + stat3
````

### Plot Gamma Probability Density
````{r}
gammaPars <- fitdistr(z$myVar,"gamma")
shapeML <- gammaPars$estimate["shape"]
rateML <- gammaPars$estimate["rate"]

stat4 <- stat_function(aes(x = xval, y = ..y..), fun = dgamma, colour="brown", n = length(z$myVar), args = list(shape=shapeML, rate=rateML))
 p1 + stat + stat2 + stat3 + stat4
````

### Plot Beta Probability Density
##### This one has to be shown in its own plot because the raw data must be rescaled so they are between 0 and 1, and then they can be compared to the beta.
```{r}
pSpecial <- ggplot(data=z, aes(x=myVar/(max(myVar + 0.1)), y=..density..)) +
  geom_histogram(color="grey60",fill="cornsilk",size=0.2) + 
  xlim(c(0,1)) +
  geom_density(size=0.75,linetype="dotted")

betaPars <- fitdistr(x=z$myVar/max(z$myVar + 0.1),start=list(shape1=1,shape2=2),"beta")
shape1ML <- betaPars$estimate["shape1"]
shape2ML <- betaPars$estimate["shape2"]

statSpecial <- stat_function(aes(x = xval, y = ..y..), fun = dbeta, colour="orchid", n = length(z$myVar), args = list(shape1=shape1ML,shape2=shape2ML))
pSpecial + statSpecial
```


#### 4. Take a look at the second-to-last graph which shows the histogram of your data and 4 probability density curves (normal, uniform, exponential, gamma) that are fit to the data. The beta distribution in the final graph is somewhat special. It often fits the data pretty well, but that is because we have assumed the largest data point is the true upper bound, and everything is scaled to that. The fit of the uniform distribution also fixes the upper bound. The other curves (normal, exponential, and gamma) are more realistic because they do not have an upper bound. For most data sets, the gamma will probably fit best, but if you data set is small, it may be very hard to see much of a difference between the curves.

#### 5. Using the best-fitting distribution, go back to the code and get the maximum likelihood parameters. Use those to simulate a new data set, with the same length as your original vector, and plot that in a histogram and add the probability density curve. Right below that, generate a fresh histogram plot of the original data, and also include the probability density curve.


### Maximum likelihood parameters for the noraml distribution
````{r}
normPars <- fitdistr(z$myVar,"normal")
print(normPars)
```


### Create simulated data and generate histogram and normal distribution prbability desnity curve
```{r}
meanML <- normPars$estimate["mean"]
sdML <- normPars$estimate["sd"]

z$new_data_normal <- rnorm(n=(length(z$ID)), mean = meanML, sd = sdML)

p2 <- ggplot(data=z, aes(x=new_data_normal, y=..density..)) +
  geom_histogram(color="grey60",fill="cornsilk",size=0.2) 
print(p2 + stat)                       
```


### Compare to new simulated data to original data histogram and probability density curve
```{r}
p1 <- ggplot(data=z, aes(x=myVar, y=..density..)) +
  geom_histogram(color="grey60",fill="cornsilk",size=0.2) 
print(p1 + stat) 
```


The simulated data and original data are similar in shape for the normal distribution, but the simulated data is less variable. 